Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-06 tex/main.tex
Start proof.
more

\begin{abstract}
SMT solvers have recently been extended with finite model finding techniques.
However, these cannot cope with universal quantifiers ranging over infinite
types. This paper introduces a translation that eliminates infinite quantifiers
in axioms specifying well-behaved recursive functions, including terminating,
tail-recursive, and productive functions. An empirical evaluation confirms that
the approach works well on benchmarks from various sources. The translation is
implemented as a preprocessor in the solver CVC4.
\end{abstract}

@@

@@

  * show that it works also for example with weird recursion schemes

@@

\ajr{Demand for recursive functions in SMT (new draft includes
define-funs-rec). Usually not interested in showing logical consistency, there
exist other tools that check e.g. termination.
Thus, want SMT solver to check satisfiability the assumption that recursive
functions are consistent (where this needs to be defined precisely).
}

\paragraph{Contributions}

In this paper, we make the following contributions:
\begin{itemize}
\item[-] We define a satisfiability-preserving translation method for a class of formulas containing recursive function definitions
that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013} are applicable.
\item[-] We provide experimental evidence that this translation improves the effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to verification conditions.
\item[-] We detail front-end support for recursive function definitions and this translation in the SMT solver \cvc.
\end{itemize}

@@

Nitpick
  * nice: three-valued logic (e.g. local overflow)
    * solving a different problem:
      * no built-in notion, what's approximated is an idealized notion, no
        injections $\concret$
  * not so nice:
    one abstract domain per type, not per function argument
  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
    odd schemes, give unsat example

>>>>

\begin{abstract}
SMT solvers have recently been extended with finite model finding techniques.
However, these cannot cope with universal quantifiers ranging over infinite
types. This paper introduces a translation that eliminates infinite quantifiers
in axioms specifying a large class of recursive functions, including
terminating, tail-recursive, and productive functions. An empirical evaluation
confirms that the approach works well on benchmarks from various sources. The
translation is implemented as a preprocessor in the solver CVC4.
\end{abstract}

@@

The recursive functions must meet a precise semantic criterion to be admissible
(Section~\ref{sec:prelim}). This criterion is met by terminating recursive
function definitions, but also by tail-recursive functions and even by
productive corecursive functions. The criterion is not met by inconsistent
definitions such as $\forall n : {\sortint}.\;\, \con{f}(n) \teq \con{f}(n) +
1$, but also by some consistent definitions.

Demand for recursive functions in SMT (new draft includes
define-funs-rec). Usually not interested in showing logical consistency, there
exist other tools that check e.g. termination.
Thus, want SMT solver to check satisfiability the assumption that recursive
functions are consistent (where this needs to be defined precisely).

We define an (un)satisfiability-preserving translation method for a class of
formulas containing recursive function definitions (Section~\ref{sec:encoding})
that produces formulas for which existing model finding techniques are
applicable. The idea

replace a definition $\forall x : \tau.\; \const{f}(x) \teq t$ by $\forall a :
\fargsort{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
$\fargsort{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
\fargsort{\tau} \to \tau$ a coercion from the abstract type to the concrete
type. For admissible definitions, this preserves satisfiability and
unsatisfiability, while making it possible to find finite models using known
techniques.

We provide experimental evidence that this translation improves the
effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to
verification conditions (Section~\ref{sec:evaluation}).

We detail front-end support for recursive function definitions and this
translation in the SMT solver \cvc (Section~\ref{sec:front-end}).

\newpage

@@

@@

@@

Nitpick
  * nice: three-valued logic (e.g. local overflow)
    * solving a different problem:
      * no built-in notion, what's approximated is an idealized notion, no
        injections $\concret$
  * not so nice:
    one abstract domain per type, not per function argument
  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
    odd schemes, give unsat example
    [* show that it works also for example with weird recursion schemes]
